---
title: "Untitled"
author: "Atmajit"
date: "June 5, 2017"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reproducible Research: Peer Assessment 1

```{r}
library(ggplot2)
library(scales)
library(Hmisc)
library(chron)
```

## Loading and preprocessing the data

1. Load the data (i.e. read.csv())

```{r pressure, echo=FALSE}
data <- read.csv('C:\\Users\\agohil\\Documents\\personal\\coursera\\reproducible\\asgnmnt1\\activity.csv')
```

## Process the data:

```{r}
total.steps <- tapply(data$steps, data$date, FUN=sum, na.rm=TRUE)
```


## What is mean total number of steps taken per day?


1. Make a histogram of the total number of steps taken each day

```{r}
library(ggplot2)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")
```

2. Calculate and report the mean and median total number of steps taken per day

```{r}
mean(total.steps, na.rm=TRUE)
median(total.steps, na.rm=TRUE)
```


## What is the average daily activity pattern?

1. Make a time series plot

```{r}
library(ggplot2)
averages <- aggregate(x=list(steps=data$steps), by=list(interval=data$interval),
                      FUN=mean, na.rm=TRUE)
ggplot(data=averages, aes(x=interval, y=steps)) +
        geom_line() +
        xlab("5-minute interval") +
        ylab("average number of steps taken")

averages[which.max(averages$steps),] 

```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
averages[which.max(averages$steps),]
```


## Imputing missing values

1. Calculate and report the total number of missing values in the dataset
```{r}
missing <- is.na(data$steps)
# How many missing
table(missing)
```

Number of missing values: r numMissingValues

There are many days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

2. Devise a strategy for filling in all of the missing values in the dataset.

To populate missing values, we choose to replace them with the mean value at the same interval across days. 

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
fill.value <- function(steps, interval) {
        filled <- NA
        if (!is.na(steps))
                filled <- c(steps)
        else
                filled <- (averages[averages$interval==interval, "steps"])
        return(filled)
}
filled.data <- data
filled.data$steps <- mapply(fill.value, filled.data$steps, filled.data$interval)

total.steps <- tapply(filled.data$steps, filled.data$date, FUN=sum)


```

4. Make a histogram of the total number of steps taken each day

```{r}
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")
```

... and Calculate and report the mean and median total number of steps taken per day.

```{r}
mean(total.steps)
median(total.steps)

```



## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two level

```{r}
library(chron)
filled.data$date <- as.Date(filled.data$date,"%m/%d/%Y")

filled.data$day <- ifelse(is.weekend(filled.data$date),"weekend","weekday")
```

2. Make a panel plot containing a time series plot

```{r}
averages <- aggregate(steps ~ interval + day, data=filled.data, mean)
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) +
        xlab("5-minute interval") + ylab("Number of steps")
```
